Plant Trait Imputation

### 2024-09-05 code/L1/L1_P5_TropicalAndes_plant_imputetraits.Rmd
minor edits to plotting and figure saving code

# Load required packages
```{r}
library(tidyr)
library(BIEN)
library(GIFT)
library(purrr)
library(mice)
library(dplyr)
library(funbiogeo)
library(visdat)
library(ggplot2)
library(forcats)
```

@@

```{r}
# Reassign NA values in TraitLevel as "Imputed"
all_traits_with_NAs$TraitLevel[is.na(all_traits_with_NAs$TraitLevel)] <- "Imputed"

# Count the number of traits per TraitLevel and TraitName
trait_counts <- all_traits_with_NAs %>%
  count(TraitLevel, TraitName) %>%
  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "Imputed"))

# Create the faceted bar plot with free y-axis scales
ggplot(trait_counts, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Trait Counts per Level",
       x = "Trait Level",
       y = "Trait Count",
       fill = "Level") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +  # Remove legend
  facet_wrap(~ TraitName, scales = "free_y") +
  guides(fill = "none")  # Remove legend
ggsave("plant_trait_counts_per_level.png", plot = last_plot(), path = figure_path)
```

```{r}
# Count the number of traits per TraitLevel
trait_counts_overall <- all_traits_with_NAs %>%
  count(TraitLevel) %>%
  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "Imputed"))

# Create the faceted bar plot with free y-axis scales
ggplot(trait_counts_overall, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Trait Counts per Level",
       x = "Trait Level",
       y = "Trait Count",
       fill = "Level") +
  theme_minimal() +
  guides(fill = "none")  # Remove legend
ggsave("plant_trait_counts_per_level_overall.png", plot = last_plot(), path = figure_path)
```

>>>>

# Load required packages
```{r}
library(tidyr)
library(BIEN)
library(GIFT)
library(purrr)
library(mice)
library(dplyr)
library(funbiogeo)
library(visdat)
library(ggplot2)
library(forcats)
library(viridis)
```

@@

```{r}
# Reassign NA values in TraitLevel as "imputed"
all_traits_with_NAs$TraitLevel[is.na(all_traits_with_NAs$TraitLevel)] <- "imputed"

# Count the number of traits per TraitLevel and TraitName
trait_counts <- all_traits_with_NAs %>%
  count(TraitLevel, TraitName) %>%
  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "imputed"))

# Create the faceted bar plot with free y-axis scales
(trait_count_level_plot <- ggplot(trait_counts, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "Trait Level",
       y = "Trait Count",
       fill = "Level") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +  # Remove legend
  facet_wrap(~ TraitName, scales = "free_y", nrow = 2) +
  guides(fill = "none"))  # Remove legend
ggsave("plant_trait_counts_per_level.png", plot = trait_count_level_plot, path = figure_path, height = 6, width = 8, units = "in")
```

```{r}
# Count the number of traits per TraitLevel
trait_counts_overall <- all_traits_with_NAs %>%
  count(TraitLevel) %>%
  mutate(TraitLevel = fct_relevel(TraitLevel, "species", "genus", "family", "imputed"))

# Create the faceted bar plot with free y-axis scales
(all_trait_count_plot <- ggplot(trait_counts_overall, aes(x = TraitLevel, y = n, fill = TraitLevel)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis(discrete = TRUE) +
  labs(title = "All traits",
       x = "Trait Level",
       y = "Trait Count",
       fill = "Level") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill = "none"))  # Remove legend
ggsave("plant_trait_counts_per_level_overall.png", plot = all_trait_count_plot, path = figure_path)
```

```{r}
(both_trait_counts <- ggarrange(trait_count_level_plot, all_trait_count_plot,
                                        ncol = 2, nrow = 1, widths = c(2,1)))
ggsave("plant_trait_counts_both.png", plot = both_trait_counts, path = figure_path, height = 6, width = 12, units = "in")
```

```{r}
# Calculate the total sum of the values
total_sum <- sum(trait_counts_overall$n)

# Calculate the percentage of each value
trait_counts_overall$percentage <- (trait_counts_overall$n / total_sum) * 100

# Print the dataframe with the percentages
print(trait_counts_overall)
```
